Semi-parametric generalized linear model for correlated binary data

### 2021-03-08 nuweb/SPGLM.tex
Remove sampling-weighted terms from gldrm
Fix bug incorporating frequency weights into initial estimate

\begin{flushleft} \small\label{scrap3}\raggedright\small
\NWtarget{nuweb3a}{} \verb@"../R/SPGLM.R"@\nobreak\ {\footnotesize {3a}}$\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@spglm_pred_mean <- function(beta, data_object, link){@\\
\mbox{}\verb@  eta <- c(data_object$model_matrix %*% beta + data_object$offset)@\\
\mbox{}\verb@  mu <- link$linkinv(eta)@\\
\mbox{}\verb@  mu@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@\\
\mbox{}\verb@spglm_loglik <- function(beta, f0, data_object, link){@\\
\mbox{}\verb@  @\\
\mbox{}\verb@    mu <- spglm_pred_mean(beta=beta, data_object=data_object, link=link)@\\
\mbox{}\verb@    N <- max(data_object$n)@\\
\mbox{}\verb@    nobs <- nrow(data_object$model_matrix)@\\
\mbox{}\verb@    @\\
\mbox{}\verb@    # ySptIndex is only used to calculate the log-likelihood, which we will not be using@\\
\mbox{}\verb@    th <- getTheta(spt=(0:N)/N, f0=f0, mu=mu, weights=data_object$weights, ySptIndex=rep(1, nobs),@\\
\mbox{}\verb@                   thetaStart=NULL, thetaControl=theta.control())@\\
\mbox{}\verb@    @\\
\mbox{}\verb@    hp <- sapply(0:N, function(t)dhyper(x=data_object$resp[,1], m=data_object$n, n=N-data_object$n, k=t))   @\\
\mbox{}\verb@    llik_term <- rowSums(t(th$fTilt) * hp)@\\
\mbox{}\verb@    llik <- llik_term %*% data_object$weights@\\
\mbox{}\verb@    c(llik)@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtFileDefBy\ \NWlink{nuweb1}{1}\NWlink{nuweb3a}{, 3a}.

@@

\begin{flushleft} \small\label{scrap7}\raggedright\small
\NWtarget{nuweb6a}{} $\langle\,${\itshape M-step}\nobreak\ {\footnotesize {6a}}$\,\rangle\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@  gldrmControl0 <- gldrm.control(returnfTiltMatrix = TRUE, returnf0ScoreInfo = FALSE, print=FALSE,@\\
\mbox{}\verb@                                betaStart = betasPre, f0Start = referencef0Pre)@\\
\mbox{}\verb@@\\
\mbox{}\verb@  mod <- gldrmFit(x = mm2, y=Ycomb[,"y"]/N, linkfun=link$linkfun, linkinv = link$linkinv,@\\
\mbox{}\verb@                  mu.eta = link$mu.eta, mu0 = mu0, offset = offset2, weights = pp, @\\
\mbox{}\verb@                  gldrmControl = gldrmControl0,  thetaControl=theta.control(),@\\
\mbox{}\verb@                  betaControl=beta.control(), f0Control=f0.control())@\\
\mbox{}\verb@                  @\\
\mbox{}\verb@  fTiltMatrix <- mod$fTiltMatrix[obs_start,]@\\
\mbox{}\verb@  betas <- mod$beta@\\
\mbox{}\verb@  referencef0 <- mod$f0@\\
\mbox{}\verb@  spt <- round(mod$spt * N)@\\
\mbox{}\verb@  llik <- c(rowSums(fTiltMatrix * hp) %*% weights)@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtMacroRefIn\ \NWlink{nuweb3b}{3b}.

@@

\begin{flushleft} \small\label{scrap8}\raggedright\small
\NWtarget{nuweb6b}{} $\langle\,${\itshape Set starting values}\nobreak\ {\footnotesize {6b}}$\,\rangle\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@  N <- max(rowSums(Y))@\\
\mbox{}\verb@  nobs <- nrow(mm)@\\
\mbox{}\verb@  p <- ncol(mm)@\\
\mbox{}\verb@  @\\
\mbox{}\verb@  betas <- NULL@\\
\mbox{}\verb@  pooled <- CBData(data.frame(Trt = "All", NResp = Y[,1], ClusterSize = rowSums(Y), Freq=ceiling(weights)), @\\
\mbox{}\verb@                    trt="Trt", clustersize="ClusterSize", nresp="NResp")@\\
\mbox{}\verb@  est <- mc.est(pooled)@\\
\mbox{}\verb@  referencef0 <- est$Prob[est$ClusterSize == N]@\\
\mbox{}\verb@  # ensure all positive values@\\
\mbox{}\verb@  referencef0 <- (referencef0 + 1e-6)/(1+(N+1)*1e-6)@\\
\mbox{}\verb@  @\\
\mbox{}\verb@  fTiltMatrix <- matrix(rep(referencef0, times=nobs), byrow=TRUE, nrow=nobs, ncol=N+1)@\\
\mbox{}\verb@  spt <- 0:N@\\
\mbox{}\verb@@\\
\mbox{}\verb@  if (is.null(mu0))@\\
\mbox{}\verb@    mu0 <- weighted.mean(Y[,1]/rowSums(Y), weights)@\\
\mbox{}\verb@  @\\
\mbox{}\verb@  # hypergeometric terms for log-likelihood calculation@\\
\mbox{}\verb@  hp <- sapply(0:N, function(t)dhyper(x=Y[,1], m=rowSums(Y), n=N-rowSums(Y), k=t))   @\\
\mbox{}\verb@  @\\
\mbox{}\verb@  # initial log-likelihood@\\
\mbox{}\verb@  llik <- rowSums(fTiltMatrix * hp) %*% weights@\\
\mbox{}\verb@@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtMacroRefIn\ \NWlink{nuweb3b}{3b}.

>>>>

\begin{flushleft} \small\label{scrap3}\raggedright\small
\NWtarget{nuweb3a}{} \verb@"../R/SPGLM.R"@\nobreak\ {\footnotesize {3a}}$\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@spglm_pred_mean <- function(beta, data_object, link){@\\
\mbox{}\verb@  eta <- c(data_object$model_matrix %*% beta + data_object$offset)@\\
\mbox{}\verb@  mu <- link$linkinv(eta)@\\
\mbox{}\verb@  mu@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@\\
\mbox{}\verb@spglm_loglik <- function(beta, f0, data_object, link){@\\
\mbox{}\verb@  @\\
\mbox{}\verb@    mu <- spglm_pred_mean(beta=beta, data_object=data_object, link=link)@\\
\mbox{}\verb@    N <- max(data_object$n)@\\
\mbox{}\verb@    nobs <- nrow(data_object$model_matrix)@\\
\mbox{}\verb@    @\\
\mbox{}\verb@    # ySptIndex is only used to calculate the log-likelihood, which we will not be using@\\
\mbox{}\verb@    th <- getTheta(spt=(0:N)/N, f0=f0, mu=mu, weights=data_object$weights, ySptIndex=rep(1, nobs),@\\
\mbox{}\verb@                   thetaStart=NULL, thetaControl=theta.control())@\\
\mbox{}\verb@    @\\
\mbox{}\verb@    hp <- sapply(0:N, function(t)dhyper(x=data_object$resp[,1], m=data_object$n, n=N-data_object$n, k=t))   @\\
\mbox{}\verb@    llik_term <- log(rowSums(t(th$fTilt) * hp))@\\
\mbox{}\verb@    llik <- llik_term %*% data_object$weights@\\
\mbox{}\verb@    c(llik)@\\
\mbox{}\verb@}@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtFileDefBy\ \NWlink{nuweb1}{1}\NWlink{nuweb3a}{, 3a}.

@@

\begin{flushleft} \small\label{scrap7}\raggedright\small
\NWtarget{nuweb6a}{} $\langle\,${\itshape M-step}\nobreak\ {\footnotesize {6a}}$\,\rangle\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@  gldrmControl0 <- gldrm.control(returnfTiltMatrix = TRUE, returnf0ScoreInfo = FALSE, print=FALSE,@\\
\mbox{}\verb@                                betaStart = betasPre, f0Start = referencef0Pre)@\\
\mbox{}\verb@@\\
\mbox{}\verb@  mod <- gldrmFit(x = mm2, y=Ycomb[,"y"]/N, linkfun=link$linkfun, linkinv = link$linkinv,@\\
\mbox{}\verb@                  mu.eta = link$mu.eta, mu0 = mu0, offset = offset2, weights = pp, @\\
\mbox{}\verb@                  gldrmControl = gldrmControl0,  thetaControl=theta.control(),@\\
\mbox{}\verb@                  betaControl=beta.control(), f0Control=f0.control())@\\
\mbox{}\verb@                  @\\
\mbox{}\verb@  fTiltMatrix <- mod$fTiltMatrix[obs_start,]@\\
\mbox{}\verb@  betas <- mod$beta@\\
\mbox{}\verb@  referencef0 <- mod$f0@\\
\mbox{}\verb@  spt <- round(mod$spt * N)@\\
\mbox{}\verb@  llik <- c(log(rowSums(fTiltMatrix * hp)) %*% weights)@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtMacroRefIn\ \NWlink{nuweb3b}{3b}.

@@

\begin{flushleft} \small\label{scrap8}\raggedright\small
\NWtarget{nuweb6b}{} $\langle\,${\itshape Set starting values}\nobreak\ {\footnotesize {6b}}$\,\rangle\equiv$
\vspace{-1ex}
\begin{list}{}{} \item
\mbox{}\verb@@\\
\mbox{}\verb@  N <- max(rowSums(Y))@\\
\mbox{}\verb@  nobs <- nrow(mm)@\\
\mbox{}\verb@  p <- ncol(mm)@\\
\mbox{}\verb@  @\\
\mbox{}\verb@  betas <- NULL@\\
\mbox{}\verb@  pooled <- CBData(data.frame(Trt = "All", NResp = Y[,1], ClusterSize = rowSums(Y), Freq=ceiling(weights)), @\\
\mbox{}\verb@                    trt="Trt", clustersize="ClusterSize", nresp="NResp", freq="Freq")@\\
\mbox{}\verb@  est <- mc.est(pooled)@\\
\mbox{}\verb@  referencef0 <- est$Prob[est$ClusterSize == N]@\\
\mbox{}\verb@  # ensure all positive values@\\
\mbox{}\verb@  referencef0 <- (referencef0 + 1e-6)/(1+(N+1)*1e-6)@\\
\mbox{}\verb@  @\\
\mbox{}\verb@  fTiltMatrix <- matrix(rep(referencef0, times=nobs), byrow=TRUE, nrow=nobs, ncol=N+1)@\\
\mbox{}\verb@  spt <- 0:N@\\
\mbox{}\verb@@\\
\mbox{}\verb@  if (is.null(mu0))@\\
\mbox{}\verb@    mu0 <- weighted.mean(Y[,1]/rowSums(Y), weights)@\\
\mbox{}\verb@  @\\
\mbox{}\verb@  # hypergeometric terms for log-likelihood calculation@\\
\mbox{}\verb@  hp <- sapply(0:N, function(t)dhyper(x=Y[,1], m=rowSums(Y), n=N-rowSums(Y), k=t))   @\\
\mbox{}\verb@  @\\
\mbox{}\verb@  # initial log-likelihood@\\
\mbox{}\verb@  llik <- log(rowSums(fTiltMatrix * hp)) %*% weights@\\
\mbox{}\verb@@\\
\mbox{}\verb@@{\NWsep}
\end{list}
\vspace{-1.5ex}
\footnotesize
\begin{list}{}{\setlength{\itemsep}{-\parsep}\setlength{\itemindent}{-\leftmargin}}
\item \NWtxtMacroRefIn\ \NWlink{nuweb3b}{3b}.
